---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(900)
```

This guide is intended as a short introduction to the fawkes package. By the end
of this guide you should know the most basic functionality of the package and be
ready to explore further with your own plots and artwork.

## Setup

To kick things off we will load the library:

```{r setup}
library(fawkes)
```

If this is your first time using fawkes you will need to install the Python 
interface to AxiDraw. This can be done very simply from the package:

```{r, eval=FALSE}
install_axidraw()
```

After that you should be good to go. One thing that you may one to do, however,
is to register the size of your AxiDraw machine. By default the Python interface 
will assume you have an A4 version and clip any commands that takes the head 
outside of that area. If you have a larger version this will result in some very
odd plots. Because this setting is something that is often set on a permanent 
basis (few of us have multiple different AxiDraw machines to switch between), it 
can be set as an option and just forget about:

```{r, eval=FALSE}
# Put this in your .Rprofile file to make it permanent
options(fawkes.model = 'A3')
```

## Package overview
fawkes comes with a bunch of different tools and operating modes. While this 
guide will mainly focus on using fawkes as a graphic device we will give a brief
overview of some of the other tools here:

### Manual control
You can control the AxiDraw completely manually, if you so desire, by creating
a manual connection with the `axi_manual()` function:

```{r, eval=FALSE}
ad <- axi_manual()
```

The returned object is an R6 object with different methods attached, such as 
`move_to()`, `pen_up()`, `line_rel()`, etc. See the documentation for 
`axi_manual()` to get a complete overview of it. The manual control is what 
drives most of the additional features in fawkes (e.g. the device capabilities).

### SVG rendering
While I would argue that the main reason for using fawkes is to use it to draw R 
graphics with it, fawkes also contains a function that takes an SVG file and 
draws it. Have a look at `axi_svg()` to get an overview of the different options
and possibilities. The functionality is a very direct wrapper around the SVG 
drawing capabilities in the Python interface. Because of this the API and 
capabilities are not comparable with the graphic device provided by fawkes. If 
you want to use the graphic device to render an SVG (e.g. to get better control 
over hatching etc) have a look at the 
[gridSVG](https://cran.r-project.org/package=gridSVG) package for rendering SVG 
through R graphics.

### Utilities
There are a couple of small utility functions that you may end up using a lot 
during the course of working with fawkes. `axi_align()` will raise the pen and 
turn the motor off, allowing you to freely move the head around (most often to
move it to the start position). `axi_pen_test()` will allow you to test 
different settings for a single pen, both when it comes to tip size, overlap, 
and different AxiDraw settings such as movement speed, delay, etc. The function
will draw a small icon showing both a filled shape and a thick line. 
`axi_pen_align()` will allow you to test different offset settings between two 
pens by overlaying an `x` and a `+` from the different pens at different 
offsets. This is an important feature if you want to mix pens with different 
barrel size.

## Using fawkes as a graphic device
fawkes comes with two different graphic devices to use. `axi_dev()` will open up
a device that continually sends instructions to the AxiDraw. While this is 
certainly a viable approach, I would at all times advice the use of the other
device: `ghost_dev()`. `ghost_dev()` opens up a device that will collect 
instructions but not send them off to the AxiDraw. These instructions can be 
inspected for correctness and plotted at your leisure. During plotting from a 
`ghost_dev()` you can pause and rewind the pen if something goes wrong which
gives a small window for error correction (pen plotters are quite unforgiving in
general). Because of these benefits this guide will focus solemnly on the 
`ghost_dev()` device.

### Some art to draw
Before we *can* draw we need something *to* draw. This should ultimately be 
something of your own doing, but for this guide we'll make something up to get
you started. We'll use the ambient package to create some noise and the isoband
package to convert it to contour lines.

```{r}
library(ambient)
library(isoband)
library(dplyr, warn.conflicts = FALSE)

noise <- long_grid(
  x = seq(-sqrt(2), sqrt(2), length.out = 1000 * sqrt(2)),
  y = seq(-1, 1, length.out = 1000)
) %>% 
  mutate(
    noise = fracture(gen_perlin, fbm, octaves = 3, x = x, y = y),
    rings = gen_waves(x = x, y = y, frequency = 0.2),
    full = noise * rings
  ) %>% 
  as.matrix(full)
```

